### 2024-10-18 rita-24/response.md
...

# Response Letter

We would like to thank the reviewers for the thoughtful comments, questions,
and criticism.

We accommodate all clarifications and answers in a final version of the paper.
To help the reviewers find the modifications of interest, we provide a link
with a *diff* between the versions, as follows:

https://

Based on the overall comments, our focus on the final version was to make the
limitations of our work more explicit, and to improve the section on related
work.

We also address each individual comment from all reviewers as follows...

- usar argumento de que centralizados precisam de bcast
- nosso caso é p2p, nem existe possibilidade de bcast pois nos nao tem
  conhecimento da rede
- possibilidade vem da existencia da time machine, que nao existe nos centralizados

In particular, the dependency on a central server

- target small P2P networks
    - examples / topology extrapolate the target

- pq exec local?
we use an absolute simulation clock to track

The topology has an average of 5 hops between peers
    - ja é bastante radical

We use NetEm [9] to simulate the network latency with
a normal distribution.

The experiments take around 40 hours to complete.

40s online and 20s offline in the
average, including peers 9-11 in the middle of the network.

-------------------------------------------------------------------------------

## Reviewer A

Regarding the experimental evaluation:

> <...>
> However, the experimental evaluation could be more extensive and rigorous.
> <...>
> The proposed approach is technically sound and the experimental results are
> promising.
> <...>
> b) The experimental evaluation could be more extensive, including more
> diverse scenarios and a comparison with existing approaches
> <...>
> The evaluation uses a fixed network topology. How would the middleware's
> performance be affected in scenarios with highly dynamic topologies, where
> peers frequently join, leave, or experience intermittent connectivity?
> Have you considered any mechanisms to adapt to such changes?
> <...>
> The evaluation focuses on simulations with up to 21 peers. How do you
> envision the middleware scaling to much larger networks, and what challenges
> might arise in real-world deployments with potentially hundreds or thousands
> of peers?

-------------------------------------------------------------------------------

Regarding the section on related work:

> a) The related work section could be expanded to include more recent work on
> peer-to-peer systems and distributed consensus algorithms;

-------------------------------------------------------------------------------

Regarding the limiations:

> <...>
> c) The paper could discuss the limitations of the proposed approach and
> potential future work.
> <...>
> The paper assumes non-malicious peers. What security measures or
> modifications would be necessary to protect against malicious peers
> attempting to disrupt the synchronization or inject false events?

-------------------------------------------------------------------------------

Regarding the middleware & protocol:

> While the API description is generally clear, some parts could be further
> clarified, particularly regarding the handling of dynamic memory allocation
> and the interaction between the application and the middleware during time
> travel.
> <...>
> Suggestion: if possible please consider providing a more formal description
> of the time synchronization protocol, including any guarantees or assumptions
> about clock drift and network delays.

## Reviewer J

Regarding the section on related work (comments J.1-J.5):

### Comment J.1

> i) This can do with a subtitles to make it easier to follow the discussion.
> There appears to be two major parts in this section: (1) the discussion on
> the three selected works (Croquet, GALS, CRDTs) in comparison to this work
> and this Work and (2) related works on time machines. It would be beneficial
> to clearly delineate these two subsection, preferably with subtitles.

As suggested, we split Section 2 in two:

- 2.1. Symmetric Distributed Applications: discussion on the selected works,
    and comparison with our proposal.
- 2.2. Software Time Machines: expanded and improved discussion on time
    machines.

### Comment J.2

> ii) The introduction on the related works makes it appear like there are
> ideally 3 subsections: (a) how the network is organized, (b) how global time
> is determined, and (c) how events are propagated and applied.
> However, this three aspects are actually related to the discussion on the
> selected works compared to the proposed solution.

Now, to make clear that we have only two subsections, we briefly introduce 2.1
and 2.2 with paragraph at the beginning of 2 (before 2.1 and 2.2):

> This section first revisits existing solutions for symmetric distributed
> applications, namely Croquet [2], GALS [1], and CRDTs [3]. Then, we discuss
> software time machines that are built for either local or networked
> applications.

We now link the three aspects (a,b,c) with the selected works at the beginning
of 2.1:

> Even though existing solutions aim to provide symmetric behavior for
> distributed applications, they diverge in three key aspects as follows:
> a. How the network is organized.
> b. How global time is determined.
> c. How events are propagated and applied.
> Figure 1 compares three selected works regarding these aspects.
> <...>

### Comment J.3

> --> There is a need to reorganize the related works section, with a proper
> introduction showing the two core aspects of the discussion, (1) and (2) as
> indicated in i), forming the introduction to the section.

Done, as detailed in J.1 and J.2.

### Comment J.4

> iii) The discussion does not have a very smooth transition. The punchline of
> the related works section should ideally be the proposal of this work, which
> is brought out more clearly when discussed at the end of the related works to
> show how all the gaps identified will be solved.

To guide our discussion, we explain how the text is organized at the beginning
of the section:

> <...>
> Figure 1 compares three selected works regarding these aspects. We now
> discuss these works with respect to the three aspects, while comparing with
> our approach. At the end of this section, we also provide initial details on
> how our middleware works.

### Comment J.5

> iv) The discussion on related work on time machines is rather abrupt.
> It does not have a proper introduction to the problem of time machines in
> relation to this work (or in general for that matter) but directly proceeds
> into an analysis of Fusion and Braid.
> --> What is aimed by this section? Answering this can help find a way of
> introducing the time machines aspect well to the reader.

The discussion was indeed abrupt and out of context, so we reshaped the new
subsection 2.2. We included two new parapgrahs at the very beginning, in which
we discuss
    (a) what a software time machine is,
    (b) in which scenarios it appears,
    (c) what are the challenges to overcome, and
    (d) what are the common implementation techniques.

-------------------------------------------------------------------------------

Regarding the protocol, middleware, and peer-to-peer concepts (comments
J.7-J.11):

### Comment J.7

> i) As a proof-of-concept, it will be important to have p2p
> model/framework to give a clear picture to an expert reader and explain to a
> lay reader the place of this middleware in the overall p2p application
> model/framework. Has this been considered?

### Comment J.8

> ii) This proposal uses broadcasting to synchronize peers. What is the
> implication of this assuming a large p2p network? Has this been considered?

The use of the term "broadcast" was inaccurate in the paper.
Now, every occurrence of "broadcast" was substituted by "trigger" or
"propagation" or "dissemination", depending on the context.

As was described in the Introduction,

> Peers in the application form a dynamic network graph and communicate only
> with direct neighbours, as in typical unstructured peer-to-peer networks.
> Events are flooded in the graph and are triggered locally with a small delay
> to compensate the network latency.

What we meant is that the same effect of a broadcast is reached after the
events are flooded in the network, but only through direct neighbour
communication.
Nevertheless, we no longer use this loosened notion of a broadcast.

We now also emphasize this distinction from centralized solutions (which do
rely on broadcasts) in the Related Work section:

> Note how the central server is the only participant with knowledge about
> clients, which never communicate directly.
> This implies that the server must never fail, and also that periodic full
> broadcasts are the only possibility of spanning the whole network.
> In this work, we argue that in a peer-to-peer alternative any node can fail,
> and that broadcasts can be replaced by optimistic flooding.
> <...>
> Note, that it is not necessary that all peers communicate directly with all
> other peers, but only indirectly via unstructured flooding.
> <...>

Section "3.2.1 Event Broadcasting" was renamed to "3.2.1 Event Dissemination"
with the introduction rephrased as follows:

> <...>, events *propagate* between peers with a timestamp scheduled to the
> future with an extra delta ∆ such that all peers are able to apply them in
> sync. To prevent *dissemination* cycles, each peer <...>

To conclude, there are no full broadcasts in the protocol.
Regardless of the network size, peers only communicate with direct neighbours.
In our experiments, flooding takes 5 hops in the average, with a maximum of 14
hops (between nodes 0 and 20).
As an example, with a *100ms* network latency, it takes *500ms* in the average
to flood an event in the P2P graph.
We detail these costs in Figure 6.
Therefore, we do consider the implication of hops and latency in the protocol.

### Comment J.9

> iii) How are the events, as well as the memory snapshots, stored for
> persistence in the p2p network, that is, is there a storage mechanisms used
> so that an offline who joins immediately picks up from where it had dropped
> off?

The full history of both events and memory snapshots are stored in all peers.
Therefore, an offline peer that joins

### Comment J.10

> iv) The simulations were done on the same machine. This means that there is
> no actual considerations due to the effect of network challenges. Is this the
> case?

### Comment J.11

> v) Although this was considering the middleware only, it would be important
> to talk about the impact in terms of network efficiency if any due to the
> middleware.

-------------------------------------------------------------------------------

Regarding grammar issues (comments J.12-J.13):

### Comment J.12

> Figure 2 - GALS --> check the spelling for client

Fixed.

### Comment J.13

> <...>
> This should be ...among all peers to be the “correct real time”,...

Fixed.